---
title: "Lab 7 Assignment: Group 21"
format:
  html:
    embed-resources: true
editor: visual
---

------------------------------------------------------------------------

# Group composition

Magdalena Zydorczak, s253712

Parastoo Tabibzadehtehrani, s252800

Sai Renuka Chandrasekaran, s252633

Ole Lennart Decker, s253728

Raquel Ruiz Pascual, s254628

------------------------------------------------------------------------

# Assignment

## Load libraries

```{r}
library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)
```

## Load data

```{r}
biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

## Data description

The biopsy dataset contains measurements from 699 breast cancer biopsies. The data include nine numerical tumor characteristics (scored from 1 to 10) and a categorical outcome indicating whether the tumor is bening or malignant.

## PCA analysis

### Look at the data in PC coordinates

### Look at the variance explained by each PC
