---
title: "Lab 7 Assignment: Group 21"
format:
  html:
    embed-resources: true
editor: visual
---

------------------------------------------------------------------------

# Group composition

Magdalena Zydorczak, s253712

Parastoo Tabibzadehtehrani, s252800

Sai Renuka Chandrasekaran, s252633

Ole Lennart Decker, s253728

Raquel Ruiz Pascual, s254628

------------------------------------------------------------------------

# Assignment

## Load libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(broom)
library(cowplot)
```

## Load data

```{r}
#| warning: false
#| message: false

biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

## Data description

The dataset analyzed in this report correponds to a collection of breast cancer biopsy measurements from 699 patients. The data include nine numerical tumor characteristics (scored from 1 to 10) and a categorical outcome indicating whether the tumor is *bening* or *malignant*.

The dataset contains nine numeric predictor variables: clump_thickness, uniform_cell_size, uniform_cell_shape, marg_adhesion, epithelial_cell_size, bare_nuclei, bland_chromatin, normal_nucleoli, mitoses, outcome.

## PCA analysis

### Look at the data in PC coordinates

```{r}
pca_fit <- biopsy |> 
  select(where(is.numeric)) |> 
  prcomp(scale = TRUE)

pca_fit |> 
  augment(biopsy) |> 
  ggplot(aes(.fittedPC1, .fittedPC2, color = outcome)) + 
  geom_point(size = 1.5, alpha=0.5) +
  scale_color_manual(
    values = c(malignant = "#9C0101", benign = "blue")
  ) +
  theme_half_open(12) + background_grid() +
  labs(title = "The first two PCs of biopsy data",
       x = "PC1",
       y = "PC2")
  
```

```{r}
arrow_style <- arrow(
  angle = 20, ends = "first", type = "closed", length = grid::unit(8, "pt")
)


pca_fit |> 
  tidy(matrix = "rotation") |> 
  pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") |> 
  ggplot(aes(PC1, PC2)) +
  geom_segment(xend = 0, yend = 0, arrow = arrow_style) +
  geom_text(
    aes(label = column),
    hjust = 1, nudge_x = -0.02, 
    color = "#9C0101"
  ) +
  xlim(-1.25, .5) + ylim(-.5, 1) +
  coord_fixed() + 
  theme_minimal_grid(12)
```

### Look at the variance explained by each PC

```{r}
pca_fit |>
  tidy(matrix = "eigenvalues") |>
  ggplot(aes(PC, percent)) +
  geom_col(fill = "#9c0101", alpha = 0.8, color = "black") +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.01))
  ) +
  theme_minimal_hgrid(12) +
  labs(y = "Percent",
       title = "Variance explained by the principal components")
```
